Simultaneous Elicitation of Scoring Rule and Agents Preferences for Robust Winner Determination

### 2020-01-20 draft.tex
some cleaning and minor fixed

In this paper we consider that both the agents’ preferences and the social choice rule are partially specified and we develop methods for computing the minimax-optimal alternative under these assumptions.
%changed from: We develop methods for computing the minimax-optimal alternative for scoring rules with partial agent preference information and partial information about the weights of the scoring rule.
While previous works have considered either partial information about the agents preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.

In this work, we focus on positional scoring rules, that are a particularly common method used to aggregate rankings and determine a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information. We then discuss several heuristics that attempt to simultaneously reduce the preference and voting rule uncertainty by determining queries, either to an agent or to the chair, that quickly reduce minimax regret.
%In particular, our query strategies focus simultaneously on reduction of relevant preference and voting rule uncertainty.

@@

Let $V$ be the set of possible preference profiles (the union, for any integer, $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
\commentBN{$s^{\profile,\w}(x)$}

where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.

@@

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} ≥ w_{2} ≥ \ldots ≥ w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
We consider now a procedure for completing a partial profile that was first proposed by \citet{Lu2011} when considering minimax regret with a fixed weight vector.
As we will show, this procedure can also be used when the weight vector is not completely known.

@@

Pick any alternative $c$ that is not $a$ or $b$. 
%Let $c_t$ denote the number of times $c$ reaches rank $t$ or better in $P$. Note that $c_m = p+q$ and $c_t ≤ p + q, 1 ≤ t ≤ m$. Define $a_t$ similarly, for the alternative $a$. By construction, $a_r = q$, $a_{r+2} = q+p = a_m$. 
Define $c'_t$ as the number of times $c$ gets rank $t$ or better in $P'$, and define $a'_t$ similarly. Now $a'_1 ≥ \delta, a'_t ≥ 2\delta \text{ for } 2 ≤ t < m, a'_{r+2} = q+p+2\delta = a'_m$. To obtain upper bounds for $c'_t$, assume $c$ is ranked first by the $p+q$ agents voting in profile $P$. By our construction, the  $2 \delta$ new agents of $P'$ give to $c$ the ranks $(3, m)$, or $(4, m-1)$, … Observe that the score of $c$ is maximal for attribution $(3, m)$, by convexity of the weights. Therefore, in the best case for $c$, $c'_1 ≤ p + q, c'_2 ≤ p+q, c'_t ≤ p+q+\delta \text{ for } 3 ≤ t < m, c'_m ≤ p+q+2\delta$. Observe that $a'_t ≥ c'_t, 1 ≤ t ≤ m$ (with strict inequality for $t=2$). Because weights are non increasing, this guarantees, with $\delta = p + q$, that the score of $a$ is not lower than the one of $c$.
\end{proof}

At this stage, we are content that a procedure exists to transform our abstract questions to questions about winners of a PSR. Further studies would investigate, for example, what is lost in terms of elicitation efficiency when we are forced to restrain to realistic concrete questions, meaning, questions involving small profiles; or investigate the relationship between our understanding of scores as (cardinal) utility and (ordinal) scores as definition of a PSR.

@@

We define here the various strategies we tested experimentally. A strategy is a function (or a random process) that gives us, given our partial knowledge so far $(\pprofile, W)$, a question that should be asked according to that strategy.

Among them, the \strat{Random} strategy is used to capture how hard an elicitation problem is, by giving us a baseline: using it we can estimate how fast the regret diminishes when we ask questions randomly. 
Choosing the questions purely randomly however, that is, independently of what we know already, would be pointless, as we would most of the time ask questions we already know the answer of. Thus, this strategy choses questions in a minimally “clever” way, just by picking questions in a set of a priori relevant questions. 
The \strat{Random} strategy first decides with a probability of $1/2$ each whether it will ask a question about weights or a question about a preference ordering (unless the profile is already known entirely). If it opted for a question about weights, it draws one rank in $2 ≤ r ≤ m-2$ equiprobably, takes the middle of the interval of values for $\lambda$ that are still possible considering our knowledge so far, and asks whether $w_r - w_{r+1} ≥ \lambda (w_{r+1} - w_{r+2})$. The intervals are initialized to $[1, n]$. If it opted for a question to agents, it draws equiprobably among the agents whose preference is not known entirely; picks an alternative $a$ randomly with equal probability among those involved in some incomparabilities in $\ppref_j$; and picks an alternative $b$ with equal probability among those incomparable with $a$ in $\ppref_j$.

% In the code, called Current solution
We also define the \strat{Extreme completions} strategy, which we consider as a more realistic baseline. 
Consider the solution of the minimax regret game, 
 $(x^{*},\bar{y},\bar{\profile},\bar{\w})$, where $x^{*}$ is the minimax regret optimal alternative, and $\bar{y} = \arg \max_{y \in A} \PMR(x,y; \pprofile, W)$, $\bar{\profile}$ and $\bar{\w}$ are the adversarial choices, thus, the instantiations that, given the choice of $x^*$, maximize the regret. 
Define: 
\begin{align}
\tau_{1} & = \min_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w),  \label{eq:tau1} \\
\tau_{2} & = \min_{\profile \in C(\pprofile)} s(\bar{y},\profile, \bar{\w}) -  s(x^{*}; \profile, \bar{\w}).
\end{align}
{\color{red}
\begin{align}
	\tau_{\text{wgt}} & = \min_{\w \in W} s^{\bar{\profile},\w}(\bar{y}) - s^{\bar{\profile},\w}(x^{*}) \\
	\tau_{i} & = \min_{\profile_i \in C(\pprofile)} s^{(\bar{\profile}_{-i},\profile_i),\bar{\w}}(\bar{y}) -  s^{(\bar{\profile}_{-i},\profile_i),\bar{\w}}(x^{*}) 
\end{align}
or we can use $\hat{\profile_i}=(\bar{\profile}_{-i},\profile_i)$ which is defined as $\bar{\profile}$ except for $i$, $\forall i \in \{1, \dots, n\}$. }

>>>>

In this paper we consider that both the agents’ preferences and the social choice rule are partially specified and we develop methods for computing the minimax-optimal alternative under these assumptions.
%changed from: We develop methods for computing the minimax-optimal alternative for scoring rules with partial agent preference information and partial information about the weights of the scoring rule.
While previous works have considered either partial information about the agents preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.
In this work, we focus on positional scoring rules, that are a particularly common method used to aggregate rankings and determine a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information. We then discuss several heuristics that attempt to simultaneously reduce the preference and voting rule uncertainty by determining queries, either to an agent or to the chair, that quickly reduce minimax regret.
%In particular, our query strategies focus simultaneously on reduction of relevant preference and voting rule uncertainty.

@@

Let $V$ be the set of possible preference profiles (the union, for any integer, $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
%\commentBN{$s^{\profile,\w}(x)$}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.

@@

We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} ≥ w_{2} ≥ \ldots ≥ w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
We consider now a procedure for completing a partial profile that was first proposed by \citet{Lu2011} when considering %minimax regret with 
a fixed weight vector.
As we will show, this procedure can also be used when the weight vector is not completely known.

@@

Pick any alternative $c$ that is not $a$ or $b$. 
%Let $c_t$ denote the number of times $c$ reaches rank $t$ or better in $P$. Note that $c_m = p+q$ and $c_t ≤ p + q, 1 ≤ t ≤ m$. Define $a_t$ similarly, for the alternative $a$. By construction, $a_r = q$, $a_{r+2} = q+p = a_m$. 
Define $c'_t$ as the number of times $c$ gets rank $t$ or better in $P'$, and define $a'_t$ similarly. Now $a'_1 ≥ \delta, a'_t ≥ 2\delta \text{ for } 2 ≤ t < m, a'_{r+2} = q+p+2\delta = a'_m$. To obtain upper bounds for $c'_t$, assume $c$ is ranked first by the $p+q$ agents voting in profile $P$. By our construction, the  $2 \delta$ new agents of $P'$ give to $c$ the ranks $(3, m)$, or $(4, m-1)$, … Observe that the score of $c$ is maximal for attribution $(3, m)$, by convexity of the weights. Therefore, in the best case for $c$, $c'_1 ≤ p + q, c'_2 ≤ p+q, c'_t ≤ p+q+\delta \text{ for } 3 ≤ t < m, c'_m ≤ p+q+2\delta$. Observe that $a'_t ≥ c'_t, 1 ≤ t ≤ m$ (with strict inequality for $t=2$). Because weights are non increasing, this guarantees, with $\delta = p + q$, that the score of $a$ is not lower than the one of $c$.
\end{proof}
At this stage, we are satisfied that a procedure exists to transform our abstract questions to questions about winners of a PSR. Further studies would investigate, for example, what is lost in terms of elicitation efficiency when we are forced to restrain to realistic concrete questions, meaning, questions involving small profiles; or investigate the relationship between our understanding of scores as (cardinal) utility and (ordinal) scores as definition of a PSR.

@@

We define here the various strategies we tested experimentally; a strategy %is a function (or a random process) that 
gives us, given the current partial knowledge $(\pprofile, W)$, a question that should be asked. % according to that strategy. 
Among them, the \strat{Random} strategy is used to capture how hard an elicitation problem is, by giving us a baseline. %: using it we can estimate how fast the regret diminishes when we ask questions randomly. 
Choosing the questions purely randomly however, that is, independently of what we know already, would be pointless, as we would most of the time ask questions we already know the answer of. Thus, this strategy choses questions in a minimally “clever” way, just by picking questions in a set of a priori relevant questions. 
The \strat{Random} strategy first decides with a probability of $1/2$ each whether it will ask a question about weights or a question about a preference ordering (unless the profile is already known entirely). If it opted for a question about weights, it draws one rank in $2 ≤ r ≤ m-2$ equiprobably, takes the middle of the interval of values for $\lambda$ that are still possible considering our knowledge so far, and asks whether $w_r - w_{r+1} ≥ \lambda (w_{r+1} - w_{r+2})$. The intervals are initialized to $[1, n]$. If it opted for a question to agents, it draws equiprobably among the agents whose preference is not known entirely; picks an alternative $a$ randomly with equal probability among those involved in some incomparabilities in $\ppref_j$; and picks an alternative $b$ with equal probability among those incomparable with $a$ in $\ppref_j$.

% In the code, called Current solution
We also define the \strat{Extreme completions} strategy, which we consider as a more realistic baseline. 
Consider the solution of the minimax regret game, 
 $(x^{*},\bar{y},\bar{\profile},\bar{\w})$, where $x^{*}$ is the minimax regret optimal alternative, and $\bar{y} = \arg \max_{y \in A} \PMR(x,y; \pprofile, W)$, $\bar{\profile}$ and $\bar{\w}$ are the adversarial choices, thus, the instantiations that, given the choice of $x^*$, maximize the regret. 
Define: 
\begin{align}
\tau_{1} & = \min_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w),  \label{eq:tau1} \\
\tau_{2} & = \min_{\profile \in C(\pprofile)} s(\bar{y},\profile, \bar{\w}) -  s(x^{*}; \profile, \bar{\w}).
\end{align}
%{\color{red}
%\begin{align}
%	\tau_{\text{wgt}} & = \min_{\w \in W} s^{\bar{\profile},\w}(\bar{y}) - s^{\bar{\profile},\w}(x^{*}) \\
%	\tau_{i} & = \min_{\profile_i \in C(\pprofile)} s^{(\bar{\profile}_{-i},\profile_i),\bar{\w}}(\bar{y}) -  s^{(\bar{\profile}_{-i},\profile_i),\bar{\w}}(x^{*}) 
%\end{align}
%or we can use $\hat{\profile_i}=(\bar{\profile}_{-i},\profile_i)$ which is defined as $\bar{\profile}$ except for $i$, $\forall i \in \{1, \dots, n\}$. }
